---
title: "Dhofar Nubian Tradition"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
    self_contained: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(kableExtra)
library(car)
library(ggsci)
library(ggtern)
library(dunn.test)
library(viridis)
library(RColorBrewer)
library(dunn.test)
library(ggpmisc)
library(broom)
library(ggtext)
library(purrr)

set.seed(132)

knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  fig.pos = "!H",
  message = FALSE,
  warning = FALSE,
  comment = NA,
  prompt = FALSE
)

options(knitr.table.format = "markdown")
options(qwraps2_markup = "markdown")
options(tinytex.verbose = TRUE)
options(warn = -1)

df = read_csv("DN_Database_20250222.csv")
df_filtered <- df %>%
  filter() %>%
  select(c("Site", "Site_Short", "Collection", "Sq_M", "Cluster_3", "Art_Type", "Art_Type_Sub", "Art_Class", "Core_Type_Tern","Condition", "Shape", "Nub_Distal_Prep", "Nub_MDR_Angle", "Scar_Pattern", "Primary_Plat", "Core_Back", "Lev_Scar", "Tool_Type", "Max_Length", "Max_Width", "Max_Thick", "Plat_Width", "Plat_Height", "Lev_Scar_Length", "Lev_Scar_Width", "Lev_Scar_Size", "TCSA", "TCSP", "Geo_Mean", "Condition", "Patination", "Dissolution", "Index_E", "Index_F", "Lev_Scar_Index_E", "Size"))

# Color Palette
virdis_colors = viridis(7, option = "D", direction = -1)
color_mapping = setNames(virdis_colors, as.character(1:7))

```
# Total assemblages

## Raw counts

__Table 1: Data table of artifact class by site__ 
```{r}
result_cluster <- df_filtered %>%
  group_by(Site, Art_Class) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(Site) %>% 
  mutate(Percentage = floor(Count / sum(Count) * 1000) / 10) %>%
  mutate(Count_Percentage = paste0(Count, " (", round(Percentage, 1), "%)")) %>%
  select(Site, Art_Class, Count_Percentage)
total_counts <- df %>%
  group_by(Site) %>%
  summarise(Total_Count = n())
pivot_result_cluster <- result_cluster %>%
  arrange(Site) %>%
  pivot_wider(
    names_from = Art_Class,
    values_from = Count_Percentage,
    values_fill = "-" 
  )
pivot_result_cluster <- pivot_result_cluster %>%
  left_join(total_counts, by = "Site")
pivot_result_cluster
```
__Figure S1: Stacked bar plots of artifact class in systematic collections__
```{r}
df_systematic <- df_filtered %>%
  filter(Collection == "systematic")
percentage_table <- df_systematic %>%
  group_by(Site, Art_Class) %>%
  summarize(count = n()) %>%
  group_by(Site) %>%
  mutate(percentage = count / sum(count) * 100)
art_class_systematic <- ggplot(percentage_table, aes(x = factor(Site), y = percentage, fill = Art_Class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  labs(x = "", y = "%", color = "Artifact Class", fill = "Artifact Class") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
art_class_systematic
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S1-art_class_systematic.png", plot = art_class_systematic, width = 12, height = 8)
```
__Figure S2: Stacked bar plots of artifact class in piece plot collections__
```{r}
df_diagnostic <- df_filtered%>%
  filter(Collection == "diagnostic")
percentage_table <- df_diagnostic %>%
  group_by(Site, Art_Class) %>%
  summarize(count = n()) %>%
  group_by(Site) %>%
  mutate(percentage = count / sum(count) * 100)
art_class_piece_plot <- ggplot(percentage_table, aes(x = factor(Site), y = percentage, fill = Art_Class)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  labs(x = "", y = "%", fill = "Artifact Class") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y  = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
art_class_piece_plot
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S2-art_class_piece_plot.png", plot = art_class_piece_plot, width = 12, height = 8)
```
## Ratios

__Figure S3: Artifact density__
```{r}
artifact_counts <- df_filtered %>%
  group_by(Site, Cluster_3) %>%
  summarize(artifact_count = n(),
            Sq_M = first(Sq_M))
artifact_counts <- artifact_counts %>%
  mutate(artifact_density = artifact_count / Sq_M)
art_density_systematic <- ggplot(artifact_counts, aes(x = Site, y = artifact_density, fill = as.factor(Cluster_3))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(artifact_density, 1)),
            vjust = -0.5, size = 3.5) +
scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  labs(title = "",
       x = "",
       y = "artifacts per sq m",
       fill = "Cluster"
       ) +
 theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
art_density_systematic
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S3-art_density_systematic.png", plot = art_density_systematic, width = 12, height = 8)
```
__Figure S4: Ratio of Lev products to prepared cores__
```{r}
artifact_counts <- df_filtered %>%
  group_by(Site, Cluster_3) %>%
  summarize(
    lev_nub_count = sum(Art_Type == "Lev_Nub", Art_Type == "Lev_cent", Art_Type == "Lev_conv"),
    point_lev_count = sum(Art_Type == "Levallois")
  ) %>%
  mutate(point_lev_to_lev_nub_ratio = ifelse(lev_nub_count == 0, NA, point_lev_count / lev_nub_count))
lev_point_ratio <- ggplot(artifact_counts, aes(x = Site, y = point_lev_to_lev_nub_ratio, fill = as.factor(Cluster_3))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = ifelse(is.na(point_lev_to_lev_nub_ratio), "NA", round(point_lev_to_lev_nub_ratio, 2))),
            vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
       x = "",
       y = "Lev products /  prepared cores",
       fill = "Cluster") +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
lev_point_ratio
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S4-lev_point_ratio.png", plot = lev_point_ratio, width = 12, height = 8)
```
__Figure S5: Ratio of retouched tools to cores__
```{r}
artifact_counts <- df_filtered %>%
  group_by(Site, Cluster_3) %>%
  summarize(
    tool = sum(Art_Class == "tool" & Tool_Type != "Levallois_blank"),
    core = sum(Art_Class == "core" & Art_Type != "indeterm")
  ) %>%
  mutate(tool_to_core_ratio = ifelse(tool == 0, NA, tool / core))
tool_core_ratio <- ggplot(artifact_counts, aes(x = Site, y = tool_to_core_ratio, fill = as.factor(Cluster_3))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = ifelse(is.na(tool_to_core_ratio), "NA", round(tool_to_core_ratio, 2))),
            vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
       x = "",
       y = "tools / cores",
       fill = "Cluster") +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
tool_core_ratio
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S5-tool_core_ratio.png", plot = tool_core_ratio, width = 12, height = 8)
```
__Figure S6: Ratio of debitage to cores__
```{r}
artifact_counts <- df_filtered %>%
  group_by(Site, Cluster_3) %>%
  summarize(
    debitage = sum(Art_Class == "debitage"),
    core = sum(Art_Class == "core")
  ) %>%
  mutate(debitage_core_ratio = ifelse(core == 0, NA, debitage / core))
debitage_core_ratio <- ggplot(artifact_counts, aes(x = Site, y = debitage_core_ratio, fill = as.factor(Cluster_3))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(debitage_core_ratio, 2)),
            vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
       x = "",
       y = "debitage / cores",
       fill = "Cluster") +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
debitage_core_ratio
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S6-debitage_core_ratio.png", plot = debitage_core_ratio, width = 12, height = 8)
```
# Patination Score Analysis

## Summary for all sites

__Weighted patination scores for each site__
```{r}
df_patination <- df_filtered %>%
  filter(!is.na(Patination))
weighted_scores <-df_filtered %>%
  group_by(Site, Patination) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Site) %>%
  summarise(
    total_count = sum(count),
    weighted_score = round(sum(Patination * count) / total_count, 1)
  )
print(weighted_scores)
```
__Figure 2: Histograms of patination scores per site__
```{r}
patina_histograms <- ggplot(data = df_filtered %>% filter(!is.na(Patination))) +
  geom_histogram(aes(y = after_stat(count), x = Patination, fill = as.factor(Patination)),
                 color = "black", binwidth = 1) +
  scale_fill_manual(values = color_mapping) +
  theme_classic() +
  scale_x_continuous(breaks = seq(1, 7, by = 1)) +
  labs(fill = "patination score", y = "artifact count", x = "patination score") +
  facet_wrap(~Site, scales = "free_y") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 13),
    axis.title.x =element_text(size = 17),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17),
    strip.text = element_text(size = 15, face = "bold")
    )
patina_histograms
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/2-patina_histograms.png", plot = patina_histograms, width = 12, height = 8)
```
__Figure S7: Stacked bar plot core type % by patination score per site__
```{r}
df_corepatina <- df_filtered %>%
  filter(Art_Class == "core", Art_Type != "pre-core&indeterm", !is.na(Patination))
percentage_table <- df_corepatina %>%
  group_by(Site, Patination, Art_Type) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
percentage_table <- percentage_table %>%
  group_by(Site, Patination) %>%
  mutate(cum_percentage = cumsum(percentage) - percentage / 2)
core_type_patination <- ggplot(percentage_table, aes(x = factor(Patination), y = percentage, fill = Art_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 2.5
  ) +
  labs(x = "patination score",
       y = "%",
      fill = "Core Type"
      ) +
  facet_wrap(~ Site) +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 17),
    legend.position = "right",
    axis.text = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17),
    strip.text = element_text(size = 15, face = "bold")
  ) +
  scale_fill_brewer(palette = "Set3")
core_type_patination
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S7-core_type_patination.png", plot = core_type_patination, width = 12, height = 8)
```
__Figure S8: Boxplots of geo mean organized by patina score__
```{r}
cores_patina_geo_mean <- ggplot(
  df_filtered %>%
    mutate(Site = factor(Site, levels = unique(Site[order(Cluster_3)]))) %>%
    filter(
      !is.na(Patination),
      Art_Class == "core",
      !is.na(Geo_Mean),
      !(Site %in% c("TH.187", "TH.188", "TH.123"))
    ),
  aes(
    x = as.factor(Patination), 
    y = Geo_Mean,
    fill = as.factor(Patination)
  )
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(
    x = "patination score",
    y = "geometric mean (mm)",
    fill = "patination score",
    title = ""
  ) +
  scale_fill_manual(values = color_mapping) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17),
    axis.title.x = element_text(size = 17),
    strip.text.x = element_text(size = 15, face = "bold")
  ) +
   facet_wrap(~ Site, scales = "fixed")
cores_patina_geo_mean
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S8-cores_patina_geo_mean_fixed.png", plot = cores_patina_geo_mean, width = 12, height = 8)
```
## Cluster Analysis

__Figure 3: Violin plot of patination scores by site with three-centroid cluster analysis__
```{r}
set.seed(132)
patination_distribution <- df_filtered %>%
  filter(!is.na(Patination)) %>%
  group_by(Site, Patination) %>%
  summarise(count = n(), .groups = 'drop') %>%
  complete(Site, Patination = 1:7, fill = list(count = 0)) %>%
  group_by(Site) %>%
  mutate(proportion = count / sum(count)) %>%
  select(-count) %>%
  pivot_wider(names_from = Patination, values_from = proportion, values_fill = list(proportion = 0))
patination_matrix <- as.matrix(patination_distribution[,-1])
kmeans_result <- kmeans(patination_matrix, centers = 3)
df_filtered <- df_filtered %>%
  mutate(Cluster = factor(kmeans_result$cluster[match(Site, patination_distribution$Site)], 
                          levels = c(3, 2, 1), labels = c("I", "II", "III")),
         Site = as.factor(Site)) %>%
  arrange(Cluster)
cluster_colors <- c("I" = "#A6CEE3", "II" = "#1F78B4", "III" = "#B2DF8A")
site_order <- df_filtered %>%
  group_by(Site) %>%
  summarize(median_patination = median(Patination, na.rm = TRUE)) %>%
  arrange(desc(median_patination)) %>%
  pull(Site)
site_order <- as.character(site_order)
index_069 <- which(site_order == "TH.069")
index_076 <- which(site_order == "TH.076")
site_order[c(index_069, index_076)] <- site_order[c(index_076, index_069)]
violin_plot <- ggplot(df_filtered, aes(x = factor(Site, levels = site_order), y = Patination, fill = Cluster)) +
  geom_violin(trim = FALSE, scale = "width") +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(breaks = seq(min(df_filtered$Patination, na.rm = TRUE), max(df_filtered$Patination, na.rm = TRUE), by = 1)) +
  theme_minimal() +
  labs(title = "",
       x = "",
       y = "Patination Score",
       fill = "Cluster") +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, vjust = 0.5, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.key = element_rect(color = "black", linewidth = 0.0)
  )
violin_plot
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/3-violin_plot.png", plot = violin_plot, width = 12, height = 8)
```
__Kruskal-Wallis and Dunn's Tests__
```{r}
kruskal_test <- kruskal.test(Patination ~ Cluster_3, data = df_filtered)
print(kruskal_test)
dunn_results <- dunn.test(df_patination$Patination, df_patination$Cluster_3, kw = TRUE, label = TRUE)
dunn_results_adjusted <- dunn_results$res
dunn_results_adjusted$Adjusted_P_Value <- p.adjust(dunn_results_adjusted$p.value, method = "bonferroni")
print(dunn_results_adjusted)
```
# All core types

## Raw counts

__Data table of core types by site__ 
```{r}
result <- df_filtered %>%
  filter(Art_Class == "core") %>%
  group_by(Site, Art_Type) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Count_Percentage = paste0(Count, " (", round(Percentage, 2), "%)")) %>%
  select(Site, Art_Type, Count_Percentage)
pivot_result <- result %>%
  arrange(Art_Type) %>%
  pivot_wider(names_from = Site, values_from = Count_Percentage, values_fill = "-") %>%
  select(Art_Type, order(as.numeric(gsub("[^0-9]", "", names(.)))))
pivot_result
```
__Data table of core types by cluster__ 
```{r}
result_cluster <- df_filtered %>%
  filter(Art_Class == "core") %>%
  group_by(Cluster_3, Art_Type) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(Cluster_3) %>% 
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Count_Percentage = paste0(Count, " (", round(Percentage, 1), "%)")) %>%
  select(Cluster_3, Art_Type, Count_Percentage)
pivot_result_cluster <- result_cluster %>%
  arrange(Art_Type) %>%
  pivot_wider(
    names_from = Cluster_3,
    values_from = Count_Percentage,
    values_fill = "-"
  ) %>%
  select(Art_Type, `I`, `II`, `III`) 
pivot_result_cluster
```
__Table 2: Data table of core types by site and cluster__ 
```{r}
process_totals_with_percentages_by_column <- function(cluster_name) {
  processed_data <- df_filtered %>%
    filter(Art_Class == "core", Cluster_3 == cluster_name) %>%
    group_by(Site, Art_Type) %>%
    summarise(Total_Count = n(), .groups = 'drop') %>%
    arrange(Art_Type) %>%
    pivot_wider(
      names_from = Site,
      values_from = Total_Count,
      values_fill = 0
    )
  processed_data_with_totals <- processed_data %>%
    rowwise() %>%
    mutate(Total = sum(c_across(starts_with("TH")), na.rm = TRUE)) %>%
    ungroup()
  processed_data_with_percentages <- processed_data_with_totals %>%
    mutate(
      across(
        .cols = starts_with("TH"),
        .fns = ~ paste0(.x, " (", round(.x / sum(.x, na.rm = TRUE) * 100, 1), "%)")
      )
    )
  final_table <- processed_data_with_percentages %>%
    select(Art_Type, everything()) %>%
    relocate(Total, .after = last_col())
  return(final_table)
}
totals_with_percentages_cluster_3_I <- process_totals_with_percentages_by_column("I")
totals_with_percentages_cluster_3_II <- process_totals_with_percentages_by_column("II")
totals_with_percentages_cluster_3_III <- process_totals_with_percentages_by_column("III")
totals_with_percentages_cluster_3_I
totals_with_percentages_cluster_3_II
totals_with_percentages_cluster_3_III
```
__Figure S9: Stacked bar plot of total core counts per site__
```{r}
df_cores <- df_filtered %>%
  filter(Art_Class == "core",
         Art_Type != "pre-core&indeterm")
count_table <- df_cores %>%
  group_by(Site, Art_Type, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop') %>%
  group_by(Site) %>%
  mutate(total_count = sum(count),
         percentage = (count / total_count) * 100)
count_table <- count_table %>%
  mutate(Site = factor(Site, levels = unique(Site)))
site_positions <- levels(count_table$Site)
total_cores_site <- ggplot(count_table %>% droplevels(), aes(x = Site, y = percentage, fill = Art_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  labs(x = "",
       y = "percentage",
      fill = "Core Type") +
  facet_wrap(~ Cluster_3, scales = "free_x") +
  theme_minimal() +
 theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17),
    strip.text = element_text(size = 15, face = "bold")
  ) +
  scale_fill_brewer(palette = "Set3")
total_cores_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S9-total_cores_site.png", plot = total_cores_site, width = 12, height = 8)
```
__Figure S10: Stacked bar plot % of total core types per cluster__
```{r}
df_cores <- df_filtered %>%
  filter(Art_Class == "core",
         Art_Type != "pre-core&indeterm")
count_table <- df_cores %>%
  group_by(Cluster_3, Art_Type) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
total_cores_cluster <- ggplot(count_table, aes(x = Cluster_3, y = percentage, fill = Art_Type)) +
  geom_bar(stat = "identity", position = "stack") +
   geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  labs(x = "cluster",
       y = "%",
       fill = "Core Type") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, hjust = 1, face = "bold"),
    axis.title = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
total_cores_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S10-total_cores_cluster.png", plot = total_cores_cluster, width = 12, height = 8)
```
## Reduction strategy summary

__Figure 5: Ternary plot of main reduction strategies by site__
```{r}
df_ternary <- df_filtered %>%
  filter(!is.na(Core_Type_Tern) & Core_Type_Tern != "")
df_counts <- df_ternary %>%
  group_by(Site_Short, Cluster_3) %>%
  summarize(
    Lev_Nub = sum(Core_Type_Tern == "Lev_Nub"),
    bi = sum(Core_Type_Tern == "bi"),
    uni = sum(Core_Type_Tern == "uni")
  )
df_proportions <- df_counts %>%
  mutate(
    total = Lev_Nub + bi + uni,
    A = Lev_Nub / total,
    B = bi / total,
    C = uni / total,
    label_x = A - .03,
    label_y = B - 0.03
  )
hulls <- df_proportions %>%
  group_by(Cluster_3) %>%
  slice(chull(A, B))
ternary_reduction <- ggtern(data = df_proportions, aes(x = A, y = B, z = C)) +
  geom_polygon(data = hulls, aes(x = A, y = B, z = C, fill = Cluster_3), alpha = 0.3, color = "black", linewidth = 0.2) +
  geom_point(aes(color = Cluster_3), size = 2.5) +
  geom_text(aes(x = label_x, y = label_y, label = Site_Short), size = 3, color = "black", fontface = "bold", hjust = -0.1) +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  theme_ggtern() +
  labs(title = "",
       x = "Nubian", y = "Bidirectional", z = "Unidirectional", color = "Cluster", fill = "Cluster") +
  theme(
    legend.position = "right",
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
ternary_reduction
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/5-ternary_reduction.png", plot = ternary_reduction, width = 12, height = 8)
```
__Chi Square Test of differences in main reduction strategies__
```{r}
contingency_table <- df_counts %>%
  select(-Cluster_3) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
__Fisher's Exact Test__
```{r}
df_counts <- df_ternary %>%
  group_by(Cluster_3, Core_Type_Tern) %>%
  summarize(Count = n(), .groups = 'drop')
df_wide <- df_counts %>%
  pivot_wider(names_from = Core_Type_Tern, values_from = Count, values_fill = list(Count = 0))
pairwise_results <- combn(unique(df_wide$Cluster_3), 2, function(pair) {
  pair_data <- df_wide %>%
    filter(Cluster_3 %in% pair) %>%
    select(-Cluster_3) %>%
    as.matrix()
  fisher_test <- fisher.test(pair_data, simulate.p.value = TRUE, B = 10000)
  list(
    clusters = pair,
    p_value = fisher_test$p.value
  )
}, simplify = FALSE)
raw_p_values <- sapply(pairwise_results, function(res) res$p_value)
adjusted_p_values <- p.adjust(raw_p_values, method = "bonferroni")
for (i in seq_along(pairwise_results)) {
  pairwise_results[[i]]$p_adj <- adjusted_p_values[i]
}
results_table <- do.call(rbind, lapply(pairwise_results, function(res) {
  data.frame(
    Cluster1 = res$clusters[1],
    Cluster2 = res$clusters[2],
    P_Value = res$p_value,
    Adjusted_P_Value = res$p_adj
  )
}))
print(results_table)
```
# Nubian Cores

## Morphology

### Median Distal Ridge (MDR)

__Figure S11: MDR angle box plots by site__
```{r}
df_mdr <- df_filtered %>%
  filter(Art_Type == "Lev_Nub", !is.na(Nub_MDR_Angle)) %>%
  arrange(Cluster_3, Site)
site_order <- df_mdr %>%
  group_by(Site) %>%
  summarize(mean_angle = mean(Nub_MDR_Angle)) %>%
  arrange(desc(mean_angle)) %>%
  pull(Site)
df_mdr$Site <- factor(df_mdr$Site, levels = site_order)
MDR_boxplot_site <- ggplot(df_mdr, aes(x = Site, y = Nub_MDR_Angle, fill = Cluster_3)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "MDR angle",
       fill = "Cluster",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1))
MDR_boxplot_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S11-MDR_angle_boxplots_site.png", plot = MDR_boxplot_site, width = 12, height = 8)
```
__Figure 7: MDR angle box plots by cluster__
```{r}
MDR_boxplot_cluster <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Nub_MDR_Angle)),
              aes(x = as.factor(Cluster_3), y = Nub_MDR_Angle,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "MDR angle",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = 13),
    axis.title.y = element_text(size = 17),
    legend.position = "none"
  )
MDR_boxplot_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/7-MDR_boxplot_cluster.png", plot = MDR_boxplot_cluster, width = 12, height = 8)
```
__MDR boxplot tests__
```{r}
df_MDR = df %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Nub_MDR_Angle))
kruskal_result <- kruskal.test(Nub_MDR_Angle ~ as.factor(Cluster_3), data = df_MDR)
print(kruskal_result)
dunn.test(df_MDR$Nub_MDR_Angle, df_MDR$Cluster_3, method = "bonferroni")
```
__Figure S12: MDR prep pie charts by cluster__
```{r}
df_mdr_p <- df_filtered %>%
  filter(Art_Type_Sub == "Lev_Nub",
         Nub_Distal_Prep != "indeterminate",
         Nub_Distal_Prep != "missing")
df_mdr_p <- df_mdr_p %>%
  mutate(Nub_Distal_Prep = factor(Nub_Distal_Prep, levels = c("divergent", "bilateral", "divergent_lateral")))
data_counts <- df_mdr_p %>%
  group_by(Cluster_3, Nub_Distal_Prep) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Nub_Distal_Prep, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
custom_labels <- c(
  "divergent" = "distal divergent (Type 1)",
  "bilateral" = "bilateral (Type 2)",
  "divergent_lateral" = "divergent-lateral (Type 1/2)"
)
MDR_prep_pie_charts <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Nub_Distal_Prep)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3", labels = custom_labels) +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.099, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
MDR_prep_pie_charts
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S12-MDR_prep_pie_charts.png", plot = MDR_prep_pie_charts, width = 12, height = 8)
```
__Chi Square Test of Nub core MDR prep between clusters__
```{r}
df_mdr_p <- df_filtered %>%
  filter(
    Art_Type_Sub == "Lev_Nub",
    Nub_Distal_Prep != "indeterminate",
    Nub_Distal_Prep != "missing",
    !is.na(Nub_Distal_Prep)
  )
contingency_table <- df_mdr_p %>%
  group_by(Cluster_3, Nub_Distal_Prep) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Nub_Distal_Prep, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
__Figure S13: Ternary plot of MDR prep by cluster__
```{r}
ddf_mdr_p <- df_filtered %>%
    filter(!is.na(Nub_Distal_Prep), Nub_Distal_Prep != "")
df_counts <- df_mdr_p %>%
    group_by(Cluster_3) %>%
summarize(
    divergent = sum(Nub_Distal_Prep == "divergent"),
        bilateral = sum(Nub_Distal_Prep == "bilateral"),
        divergent_lateral = sum(Nub_Distal_Prep == "divergent_lateral")
    )
df_proportions <- df_counts %>%
    mutate(
       total = divergent + bilateral + divergent_lateral,
       total = ifelse(total == 0, 1, total),
       A = divergent / total,
       B = bilateral / total,
       C = divergent_lateral / total,
      label_x = A - .02,
       label_y = B + 0.02
      )
MDR_prep_tern <- ggtern(data = df_proportions, aes(x = A, y = B, z = C)) +
  geom_point(aes(color = Cluster_3), size = 2.5) +
  geom_text(aes(x = label_x, y = label_y, label = Cluster_3), size = 3, color = "black", fontface = "bold", hjust = -0.1) +
  scale_color_brewer(palette = "Paired") +
  theme_ggtern() +
  labs(title = "",
       x = "Type 1", y = "Type 2", z = "Type 1/2", color = "Cluster") +
  theme(
    legend.position = "right",
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  )
MDR_prep_tern
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S13-MDR_prep_tern.png", plot = MDR_prep_tern, width = 12, height = 8)
```
### Platforms

__Figure S14: Nubian platform prep pie charts for all sites__
```{r}
df_nub_plat <- df_filtered %>%
  filter(Art_Type_Sub == "Lev_Nub",
         Primary_Plat != "crushed&missing&indeterm",
         !is.na(Primary_Plat))
data_counts <- df_nub_plat %>%
  group_by(Site, Primary_Plat, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Primary_Plat, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
nub_plat_prep_site <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x", ncol = 5) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
nub_plat_prep_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S14-nub_plat_prep_site.png", plot = nub_plat_prep_site, width = 12, height = 8)
```
__Figure S15: Nubian platform prep pie charts by cluster__
```{r}
df_nub_plat <- df_filtered %>%
  filter(Art_Type_Sub == "Lev_Nub", Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_nub_plat %>%
  group_by(Cluster_3, Primary_Plat) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Primary_Plat, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
nub_plat_prep_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
   theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
nub_plat_prep_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S15-nub_plat_prep_cluster.png", plot = nub_plat_prep_cluster, width = 12, height = 8)
```
### Shape

__Figure S16: Nubian core shape pie charts for all sites__
```{r}
df_nub_s <- df_filtered %>%
  filter(Art_Type_Sub == "Lev_Nub", !is.na(Shape), Shape != "other", Shape != "irregular")
data_counts <- df_nub_s %>%
  group_by(Site, Shape, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Shape, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
nub_shape_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
nub_shape_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S16-nub_shape_sites.png", plot = nub_shape_sites, width = 12, height = 8)
```
__Figure S17: Nubian core shape pie charts by cluster__
```{r}
df_nub_s <- df_filtered %>%
  filter(Art_Type_Sub == "Lev_Nub", !is.na(Shape), Shape != "indeterminate", Shape != "irregular", Shape != "other")
data_counts <- df_nub_s %>%
  group_by(Cluster_3, Shape) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Shape, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
Nub_core_shape_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
Nub_core_shape_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S17-nub_core_shape_cluster.png", plot = Nub_core_shape_cluster, width = 12, height = 8)
```
__Chi Square Test of Nub core shape between clusters__
```{r}
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Shape, count) %>%
  pivot_wider(names_from = Shape, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
__Fisher's Exact Test__
```{r}
clusters <- unique(as.character(data_counts_complete$Cluster_3))
cluster_pairs <- combn(clusters, 2, simplify = FALSE)
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Shape, count) %>%
    spread(key = Shape, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  subset_table <- subset_table[rowSums(subset_table) > 0, colSums(subset_table) > 0]
  if (nrow(subset_table) > 1 && ncol(subset_table) > 1) {
    fisher_result <- fisher.test(subset_table, simulate.p.value = TRUE)
    p_value <- fisher_result$p.value
  } else {
    p_value <- NA
  }
  tibble(
    Cluster_1 = cluster1,
    Cluster_2 = cluster2,
    Fisher_p_value = p_value
  )
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
posthoc_results_fisher <- posthoc_results_fisher %>%
  mutate(p_value_adjusted = p.adjust(Fisher_p_value, method = "bonferroni"))
print(posthoc_results_fisher)
```
### Final Observable Scar

__Figure S18: Nubian core final scar for all sites__
```{r}
df_nub_fls <- df_filtered %>%
  filter(Art_Type == "Lev_Nub", !is.na(Lev_Scar), Lev_Scar != "recycled", Lev_Scar != "overpassed")
data_counts <- df_nub_fls %>%
  group_by(Site, Lev_Scar, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Lev_Scar, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
nub_fls_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Lev_Scar)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
 labs(title = "", fill = "") +  
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
nub_fls_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S18-nub_fls_sites.png", plot = nub_fls_sites, width = 12, height = 8)
```
__Figure S19: Nubian core final scar pie charts by cluster__
```{r}
df_nub_fls <- df_filtered %>%
  filter(Art_Type == "Lev_Nub", !is.na(Lev_Scar), Lev_Scar != "recycled", Lev_Scar != "overpassed")
data_counts <- df_nub_fls %>%
  group_by(Cluster_3, Lev_Scar) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Lev_Scar, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
Nub_core_fls_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Lev_Scar)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
Nub_core_fls_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S19-nub_core_fls_cluster.png", plot = Nub_core_fls_cluster, width = 12, height = 8)
```
__Chi Square Test of Nub core final scar between clusters__
```{r}
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Lev_Scar, count) %>%
  pivot_wider(names_from = Lev_Scar, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
## Metric Analysis

### Linear Measurements

__Figure S20: Boxplot of Nub core length by site__
```{r}
Nub_core_length_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Length)),
  aes(x = reorder(Site, desc(Max_Length), FUN = median), y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S20-nub_core_length_site.png", plot = Nub_core_length_site, width = 12, height = 8)
```
__Figure S21: Boxplot of Nub core length by cluster__
```{r}
Nub_core_length_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Length), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Max_Length,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max length (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S21-nub_core_length_cluster.png", plot = Nub_core_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis Nub Core Length Test__
```{r}
df_Nub_L = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Length))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_Nub_L)
dunn.test(df_Nub_L$Max_Length, df_Nub_L$Cluster_3, method = "bonferroni")
```
__Figure S22: Boxplot of Nub core width by site__
```{r}
Nub_core_width_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Width)),
  aes(x = reorder(Site, desc(Max_Width), FUN = median), y = Max_Width, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S22-nub_core_width_site.png", plot = Nub_core_width_site, width = 12, height = 8)
```
__Figure S23: Boxplot of Nub core width by cluster__
```{r}
Nub_core_width_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Width), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Max_Width,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S23-nub_core_width_cluster.png", plot = Nub_core_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis Nub Core Width Test__
```{r}
df_Nub_W = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Width))
kruskal.test(Max_Width ~ as.factor(Cluster_3), data = df_Nub_W)
dunn.test(df_Nub_W$Max_Width, df_Nub_W$Cluster_3, method = "bonferroni")
```
__Figure S24: Boxplot of Nubian core thickness by site__
```{r}
Nub_core_thick_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Thick)),
  aes(x = reorder(Site, desc(Max_Thick), FUN = median), y = Max_Thick, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickness (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_thick_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S24-nub_core_thick_site.png", plot = Nub_core_thick_site, width = 12, height = 8)
```
__Figure S25: Boxplot of Nub core thickness by cluster__
```{r}
Nub_core_thick_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Thick), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Max_Thick,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "midpoint thickness (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_thick_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S25-nub_core_thick_cluster.png", plot = Nub_core_thick_cluster, width = 12, height = 8)
```
__Kruskal Wallis Nub Core Thick Test__
```{r}
df_Nub_T = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Thick))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_Nub_T)
dunn.test(df_Nub_T$Max_Thick, df_Nub_T$Cluster_3, method = "bonferroni")
```
__Figure S26: Boxplot of Nubian core Index of Flattening by site__
```{r}
Nub_core_IF_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Index_F)),
  aes(x = reorder(Site, desc(Index_F), FUN = median), y = Index_F, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_IF_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S26-nub_core_IF_site.png", plot = Nub_core_IF_site, width = 12, height = 8)
```
__Figure S27: Boxplot of Nubian core Index of Flattening by cluster__
```{r}
Nub_core_IF_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_F), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Index_F,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "thickness / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_IF_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S27-nub_core_IF_cluster.png", plot = Nub_core_IF_cluster, width = 12, height = 8)
```
__Kruskal Wallis Test of Nub Core IF__
```{r}
df_Nub_IF = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Index_F))
kruskal.test(Index_F ~ as.factor(Cluster_3), data = df_Nub_IF)
dunn.test(df_Nub_IF$Index_F, df_Nub_IF$Cluster_3, method = "bonferroni")
```
__Figure S28: Boxplot of Nubian core Index of Elongation by site__
```{r}
Nub_core_IE_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Index_E)),
  aes(x = reorder(Site, desc(Index_E), FUN = median), y = Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S28-nub_core_IE_site.png", plot = Nub_core_IE_site, width = 12, height = 8)
```
__Figure S29: Boxplot of Nubian core Index of Elongation by cluster__
```{r}
Nub_core_IE_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_E), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S29-nub_core_IE_cluster.png", plot = Nub_core_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Tests of Nub Core IE__
```{r}
df_Nub_IE = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Index_E))
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_Nub_IE)
dunn.test(df_Nub_IE$Index_E, df_Nub_IE$Cluster_3, method = "bonferroni")
```
__Figure S30: Nub core L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
Nub_core_LW_scatter <- ggplot(
    df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 1.1, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 30, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
Nub_core_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S30-nub_core_LW_scatter.png", plot = Nub_core_LW_scatter, width = 12, height = 8)
```
__Figure 8: Boxplot of Nubian core geometric mean by site__
```{r}
Nub_core_geo_mean_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Geo_Mean)),
  aes(x = reorder(Site, desc(Geo_Mean), FUN = median), y = Geo_Mean, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
Nub_core_geo_mean_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/8-nub_core_geo_mean_site.png", plot = Nub_core_geo_mean_site, width = 12, height = 8)
```
__Figure S31: Boxplot of Nubian core geometric mean by cluster__
```{r}
Nub_core_geo_mean_cluster <- ggplot(df_filtered %>% filter(!is.na(Geo_Mean), Art_Type_Sub == "Lev_Nub"),
              aes(x = as.factor(Cluster_3), y = Geo_Mean,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "geometric mean (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
Nub_core_geo_mean_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S31-nub_core_geo_mean_cluster.png", plot = Nub_core_geo_mean_cluster, width = 12, height = 8)
```
__Kruskal Wallis & Dunns Tests of Nub core geo mean__
```{r}
df_NGM = df_filtered %>% filter(Art_Type_Sub == "Lev_Nub", !is.na(Geo_Mean))
kruskal.test(Geo_Mean ~ as.factor(Cluster_3), data = df_NGM)
dunn.test(df_NGM$Geo_Mean, df_NGM$Cluster_3, method = "bonferroni")
```
# Bidirectional Cores

## Morphology

### Dorsal Scars

__Figure S32: Bi core dorsal scar pie charts for all sites__
```{r}
df_bi_ds <- df_filtered %>%
  filter(Art_Type == "bi", !is.na(Scar_Pattern))
data_counts <- df_bi_ds %>%
  group_by(Site, Scar_Pattern, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Scar_Pattern, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
bi_ds_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Scar_Pattern)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
bi_ds_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S32-bi_ds_sites.png", plot = bi_ds_sites, width = 12, height = 8)
```
__Figure S33: Bi core shape pie charts by cluster__
```{r}
df_bi_ds <- df_filtered %>%
  filter(Art_Type == "bi", !is.na(Scar_Pattern))
data_counts <- df_bi_ds %>%
  group_by(Cluster_3, Scar_Pattern) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Scar_Pattern, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
bi_ds_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Scar_Pattern)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
bi_ds_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S33-bi_core_ds_cluster.png", plot = bi_ds_cluster, width = 12, height = 8)
```
__Chi Square Test of Bi core dorsal scars between clusters__
```{r}
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Scar_Pattern, count) %>%
  pivot_wider(names_from = Scar_Pattern, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
__Fisher's Exact Test__
```{r}
clusters <- unique(as.character(data_counts_complete$Cluster_3))
cluster_pairs <- combn(clusters, 2, simplify = FALSE)
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Scar_Pattern, count) %>%
    spread(key = Scar_Pattern, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  subset_table <- subset_table[rowSums(subset_table) > 0, colSums(subset_table) > 0]
  if (nrow(subset_table) > 1 && ncol(subset_table) > 1) {
    fisher_result <- fisher.test(subset_table, simulate.p.value = TRUE)
    p_value <- fisher_result$p.value
  } else {
    p_value <- NA
  }
  tibble(
    Cluster_1 = cluster1,
    Cluster_2 = cluster2,
    Fisher_p_value = p_value
  )
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
posthoc_results_fisher <- posthoc_results_fisher %>%
  mutate(p_value_adjusted = p.adjust(Fisher_p_value, method = "bonferroni"))
print(posthoc_results_fisher)
```
### Platforms

__Figure S34: Bi core platform prep pie charts for all sites__
```{r}
df_bi_p <- df_filtered %>%
  filter(Art_Type == "bi", Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_bi_p %>%
  group_by(Site, Primary_Plat, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Primary_Plat, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
bi_plat_prep_site <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x", ncol = 5) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 2.5
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
bi_plat_prep_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S34-bi_plat_prep_site.png", plot = bi_plat_prep_site, width = 12, height = 8)
```
__Figure S35: Bi core platform prep pie charts by cluster__
```{r}
df_bi_p <- df_filtered %>%
  filter(Art_Type == "bi", Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_bi_p %>%
  group_by(Cluster_3, Primary_Plat) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Primary_Plat, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
bi_core_plat <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 12),   
    legend.position = "bottom",
    strip.text = element_text(size = 14, face = "bold")
  )
bi_core_plat
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S35-bi_core_plat.png", plot = bi_core_plat, width = 12, height = 8)
```
__Chi Square Test of Platform differences between clusters__
```{r}
df_bi_p <- df_filtered %>%
  filter(Art_Type == "bi", Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts_bi <- df_bi_p %>%
  group_by(Cluster_3, Primary_Plat) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_bi_complete <- data_counts_bi %>%
  complete(Cluster_3, Primary_Plat, fill = list(count = 0))
contingency_table <- data_counts_bi_complete %>%
  select(Cluster_3, Primary_Plat, count) %>%
  pivot_wider(names_from = Primary_Plat, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
### Shape

__Figure S36: Bi core shape pie charts for all sites__
```{r}
df_bi_s <- df_filtered %>%
  filter(Art_Type == "bi", !is.na(Shape), Shape != "other", Shape != "indeterminate")
data_counts <- df_bi_s %>%
  group_by(Site, Shape, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Shape, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
bi_shape_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
bi_shape_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S36-bi_shape_sites.png", plot = bi_shape_sites, width = 12, height = 8)
```
__Figure S37: Bi core shape pie charts by cluster__
```{r}
df_bi_s <- df_filtered %>%
  filter(Art_Type == "bi", Shape != "other", !is.na(Shape))
data_counts <- df_bi_s %>%
  group_by(Cluster_3, Shape) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Shape, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
custom_labels <- c(
  "cordiform" = "cordiform",
  "cylindrical" = "cylindrical",
  "nws" = "narrow working surface",
  "rectangular" = "rectangular",
  "triangular" = "triangular"
)
bi_core_shape <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_manual(values = RColorBrewer::brewer.pal(n = length(custom_labels), name = "Set3"),
                    labels = custom_labels)  +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
bi_core_shape
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S37-bi_core_shape_cluster.png", plot = bi_core_shape, width = 12, height = 8)
```
__Chi Square test of shape differences between clusters__
```{r}
df_bi_s <- df_filtered %>%
  filter(Art_Type == "bi", Shape != "other", !is.na(Shape))
data_counts_bi <- df_bi_s %>%
  group_by(Cluster_3, Shape) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_bi_complete <- data_counts_bi %>%
  complete(Cluster_3, Shape, fill = list(count = 0))
contingency_table <- data_counts_bi_complete %>%
  select(Cluster_3, Shape, count) %>%
  spread(key = Shape, value = count, fill = 0) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
chisq_test
```
__Fisher's Exact Test__
```{r}
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_bi_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Shape, count) %>%
    spread(key = Shape, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  if (nrow(subset_table) != 2 || ncol(subset_table) < 2) {
    warning(paste("Skipping invalid table for pair:", cluster1, cluster2))
    return(tibble(
      Cluster_1 = cluster1,
      Cluster_2 = cluster2,
      Fisher_p_value = NA,
      p_value_adjusted = NA
    ))
  }
  fisher_result <- fisher.test(subset_table, simulate.p.value = TRUE, B = 1e5)
  tibble(
    Cluster_1 = cluster1,
    Cluster_2 = cluster2,
    Fisher_p_value = fisher_result$p.value,
    p_value_adjusted = p.adjust(fisher_result$p.value, method = "bonferroni")
  )
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
print(posthoc_results_fisher)
```
### Final Observable Scar

__Figure S38: Bi core final scar pie charts for all sites__
```{r}
df_bi_fls <- df_filtered %>%
  filter(Art_Type == "bi", !is.na(Lev_Scar), Lev_Scar != "recycled", Lev_Scar != "overpassed")
data_counts <- df_bi_fls %>%
  group_by(Site, Lev_Scar, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Lev_Scar, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
bi_fls_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Lev_Scar)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
 labs(title = "", fill = "") +  
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
bi_fls_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S38-bi_fls_sites.png", plot = bi_fls_sites, width = 12, height = 8)
```
__Figure S39: Bi core final scar pie charts by cluster__
```{r}
df_bi_fls <- df_filtered %>%
  filter(Art_Type == "Lev_Nub", !is.na(Lev_Scar), Lev_Scar != "recycled", Lev_Scar != "overpassed")
data_counts <- df_bi_fls %>%
  group_by(Cluster_3, Lev_Scar) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Lev_Scar, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
Nub_core_fls_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Lev_Scar)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
Nub_core_fls_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S39-nub_core_fls_cluster.png", plot = Nub_core_fls_cluster, width = 12, height = 8)
```
__Chi Square Test of bi core final scar between clusters__
```{r}
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Lev_Scar, count) %>%
  pivot_wider(names_from = Lev_Scar, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
## Metric Analysis

__Figure S40: Boxplot of bi core length by site__
```{r}
bi_core_length_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Length)),
  aes(x = reorder(Site, desc(Max_Length), FUN = median), y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S40-bi_core_length_site.png", plot = bi_core_length_site, width = 12, height = 8)
```
__Figure S41: Boxplot of Nub core length by cluster__
```{r}
bi_core_length_cluster <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Length)),
              aes(x = as.factor(Cluster_3), y = Max_Length,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max length (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S41-bi_core_length_cluster.png", plot = bi_core_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Test of bi core length__
```{r}
df_bi_L = df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Length))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_bi_L)
dunn.test(df_bi_L$Max_Length, df_bi_L$Cluster_3, method = "bonferroni")
```
__Figure S42: Boxplot of bi core width by site__
```{r}
bi_core_width_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Width)),
  aes(x = reorder(Site, desc(Max_Width), FUN = median), y = Max_Width, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S42-bi_core_width_site.png", plot = bi_core_width_site, width = 12, height = 8)
```
__Figure S43: Boxplot of bi core width by cluster__
```{r}
bi_core_width_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Width), Art_Type == "bi"),
              aes(x = as.factor(Cluster_3), y = Max_Width,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S43-bi_core_width_cluster.png", plot = bi_core_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of bi core width__
```{r}
df_bi_W = df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Width))
kruskal.test(Max_Width ~ as.factor(Cluster_3), data = df_bi_W)
dunn.test(df_bi_W$Max_Width, df_bi_W$Cluster_3, method = "bonferroni")
```
__Figure S44: Boxplot of bi core thickness by site__
```{r}
bi_core_thick_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Thick)),
  aes(x = reorder(Site, desc(Max_Thick), FUN = median), y = Max_Thick, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickness (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_thick_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S44-bi_core_thick_site.png", plot = bi_core_thick_site, width = 12, height = 8)
```
__Figure S45: Boxplot of bi core thickness by cluster__
```{r}
bi_core_thick_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Thick), Art_Type == "bi"),
              aes(x = as.factor(Cluster_3), y = Max_Thick,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "midpoint thickness (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_thick_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S45-bi_core_thick_cluster.png", plot = bi_core_thick_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of bi core thickness__
```{r}
df_bi_T = df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Thick))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_bi_T)
dunn.test(df_bi_T$Max_Thick, df_bi_T$Cluster_3, method = "bonferroni")
```
__Figure S46: Boxplot of bi core Index of Flattening by site__
```{r}
bi_core_IF_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Index_F)),
  aes(x = reorder(Site, desc(Index_F), FUN = median), y = Index_F, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_IF_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S46-bi_core_IF_site.png", plot = bi_core_IF_site, width = 12, height = 8)
```
__Figure S47: Boxplot of bi core Index of Flattening by cluster__
```{r}
bi_core_IF_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_F), Art_Type == "bi"),
              aes(x = as.factor(Cluster_3), y = Index_F,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "thickness / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_IF_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S47-bi_core_IF_cluster.png", plot = bi_core_IF_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of bi core IF__
```{r}
df_bi_IF = df_filtered %>% filter(Art_Type == "bi", !is.na(Index_F))
kruskal.test(Index_F ~ as.factor(Cluster_3), data = df_bi_IF)
dunn.test(df_bi_IF$Index_F, df_bi_IF$Cluster_3, method = "bonferroni")
```
__Figure S48: Boxplot of bi core Index of Elongation by site__
```{r}
bi_core_IE_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Index_E)),
  aes(x = reorder(Site, desc(Index_E), FUN = median), y = Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S48-bi_core_IE_site.png", plot = bi_core_IE_site, width = 12, height = 8)
```
__Figure S49: Boxplot of bi core Index of Elongation by cluster__
```{r}
bi_core_IE_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_E), Art_Type == "bi"),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S49-bi_core_IE_cluster.png", plot = bi_core_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of bi core IE__
```{r}
df_bi_IE = df_filtered %>% filter(Art_Type == "bi", !is.na(Index_E))
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_bi_IE)
```
__Figure S50: bi core L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Type == "bi", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(Art_Type == "bi", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
bi_core_LW_scatter <- ggplot(
    df_filtered %>% filter(Art_Type == "bi", !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 1.1, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 30, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
bi_core_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S50-bi_core_LW_scatter.png", plot = bi_core_LW_scatter, width = 12, height = 8)
```
__Figure 11: Boxplot of bi core geometric mean by site__
```{r}
bi_core_geo_mean_site <- ggplot(df_filtered %>% filter(Art_Type == "bi", !is.na(Geo_Mean)),
  aes(x = reorder(Site, desc(Geo_Mean), FUN = median), y = Geo_Mean, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
bi_core_geo_mean_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/11-bi_core_geo_mean_site.png", plot = bi_core_geo_mean_site, width = 12, height = 8)
```
__Figure S51: Boxplot of bi core geometric mean by cluster__
```{r}
bi_core_geo_mean_cluster <- ggplot(df_filtered %>% filter(!is.na(Geo_Mean), Art_Type == "bi"),
              aes(x = as.factor(Cluster_3), y = Geo_Mean,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "geometric mean (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
bi_core_geo_mean_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S51-bi_core_geo_mean_cluster.png", plot = bi_core_geo_mean_cluster, width = 12, height = 8)
```
__Kruskal Wallis & Dunn's Tests of bi core geo mean__
```{r}
df_BGM = df_filtered %>% filter(Art_Type == "bi", !is.na(Geo_Mean))
kruskal.test(Geo_Mean ~ as.factor(Cluster_3), data = df_BGM)
dunn.test(df_BGM$Geo_Mean, df_BGM$Cluster_3, method = "bonferroni")
```
# Unidirectional cores

## Morphology

### Dorsal Scars

__Figure S52: Uni core dorsal scar pie charts for all sites__
```{r}
df_uni_ds <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Scar_Pattern))
data_counts <- df_uni_ds %>%
  group_by(Site, Scar_Pattern, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Scar_Pattern, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
uni_ds_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Scar_Pattern)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
uni_ds_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S52-uni_ds_sites.png", plot = uni_ds_sites, width = 12, height = 8)
```
__Figure S53: Uni core dorsal scar pie charts by cluster__
```{r}
df_uni_ds <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Scar_Pattern))
data_counts <- df_uni_ds %>%
  group_by(Cluster_3, Scar_Pattern) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Scar_Pattern, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
uni_ds_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Scar_Pattern)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
uni_ds_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S53-uni_core_ds_cluster.png", plot = uni_ds_cluster, width = 12, height = 8)
```
__Chi Square Test of uni core dorsal scars between clusters__
```{r}
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Scar_Pattern, count) %>%
  pivot_wider(names_from = Scar_Pattern, values_from = count, values_fill = list(count = 0)) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
print(chisq_test)
```
__Fisher's Exact Test__
```{r}
clusters <- unique(as.character(data_counts_complete$Cluster_3))
cluster_pairs <- combn(clusters, 2, simplify = FALSE)
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Scar_Pattern, count) %>%
    spread(key = Scar_Pattern, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  subset_table <- subset_table[rowSums(subset_table) > 0, colSums(subset_table) > 0]
  if (nrow(subset_table) > 1 && ncol(subset_table) > 1) {
    fisher_result <- fisher.test(subset_table, simulate.p.value = TRUE)
    p_value <- fisher_result$p.value
  } else {
    p_value <- NA
  }
  tibble(
    Cluster_1 = cluster1,
    Cluster_2 = cluster2,
    Fisher_p_value = p_value
  )
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
posthoc_results_fisher <- posthoc_results_fisher %>%
  mutate(p_value_adjusted = p.adjust(Fisher_p_value, method = "bonferroni"))
print(posthoc_results_fisher)
```
### Shape

__Figure S54: Uni core platform prep pie charts for all sites__
```{r}
df_uni_p <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_uni_p %>%
  group_by(Site, Primary_Plat, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Primary_Plat, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
uni_plat_prep_site <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x", ncol = 5) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 2.5
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
uni_plat_prep_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S54-uni_plat_prep_site.png", plot = uni_plat_prep_site, width = 12, height = 8)
```
__Figure S55: Uni core platform prep pie charts by cluster__
```{r}
df_uni_p <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_uni_p %>%
  group_by(Cluster_3, Primary_Plat) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Primary_Plat, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
uni_core_plat_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Primary_Plat)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
uni_core_plat_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S55-uni_core_plat_cluster.png", plot = uni_core_plat_cluster, width = 12, height = 8)
```
__Chi Square Test of Platform differences between clusters__
```{r}
df_uni_p <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), Primary_Plat != "crushed&missing&indeterm", !is.na(Primary_Plat))
data_counts <- df_uni_p %>%
  group_by(Cluster_3, Primary_Plat) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Primary_Plat, fill = list(count = 0))
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Primary_Plat, count) %>%
  spread(key = Primary_Plat, value = count, fill = 0) %>%
  column_to_rownames(var = "Cluster_3")
print(contingency_table)
chisq_test <- chisq.test(contingency_table)
chisq_test
```
__Figure S56: Uni core shape pie charts for all sites__
```{r}
df_uni_s <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Shape), Shape != "indeterminate", Shape != "irregular")
data_counts <- df_uni_s %>%
  group_by(Site, Shape, Cluster_3) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Site, Shape, Cluster_3, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Site, Cluster_3) %>%
  mutate(proportion = count / sum(count)) %>%
  filter(sum(count) > 0)
data_counts_filtered <- data_counts_complete %>%
  droplevels() %>%
  filter(Site %in% unique(Site[!is.na(Cluster_3)]))
uni_shape_sites <- ggplot(data_counts_filtered, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3 + Site, scales = "free_x") +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "", fill = "") +
   geom_text(
    aes(label = ifelse(proportion > 0.089, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold")
  )
uni_shape_sites
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S56-uni_shape_sites.png", plot = uni_shape_sites, width = 12, height = 8)
```
__Figure S57: uni core shape pie charts by cluster__
```{r}
df_uni_s <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Shape), Shape != "irregular", Shape != "other", Shape != "indeterminate")
data_counts <- df_uni_s %>%
  group_by(Cluster_3, Shape) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Shape, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
custom_labels <- c(
  "cordiform" = "cordiform",
  "cylindrical" = "cylindrical",
  "nws" = "narrow working surface",
  "other" = "other",
  "rectangular" = "rectangular",
  "triangular" = "triangular"
)
uni_core_shape <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Shape)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
 scale_fill_manual(values = brewer.pal(n = length(custom_labels), "Set3"),
                    labels = custom_labels) +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
uni_core_shape
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S57-uni_core_shape.png", plot = uni_core_shape, width = 12, height = 8)
```
__Chi Square Test of shape differences between clusters__
```{r}
df_uni_s <- df_filtered %>%
  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Shape), Shape != "irregular", Shape != "other", Shape != "indeterminate")
data_counts <- df_uni_s %>%
  group_by(Cluster_3, Shape) %>%
  summarize(count = n(), .groups = 'drop') %>%
  mutate(Cluster_3 = as.factor(Cluster_3), Shape = as.factor(Shape))
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Shape, fill = list(count = 0))
print(contingency_table)
contingency_table <- data_counts_complete %>%
  select(Cluster_3, Shape, count) %>%
  spread(key = Shape, value = count, fill = 0) %>%
  column_to_rownames(var = "Cluster_3")
contingency_table <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
chisq_test <- chisq.test(contingency_table)
chisq_test
```
__Fisher's Exact Test__
```{r}
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Shape, count) %>%
    spread(key = Shape, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  subset_table <- subset_table[rowSums(subset_table) > 0, colSums(subset_table) > 0]
  if (nrow(subset_table) > 1 && ncol(subset_table) > 1) {
    fisher_result <- fisher.test(subset_table, simulate.p.value = TRUE, B = 1e5)
    tibble(
      Cluster_1 = cluster1,
      Cluster_2 = cluster2,
      Fisher_p_value = fisher_result$p.value
    )
  } else {
    tibble(
      Cluster_1 = cluster1,
      Cluster_2 = cluster2,
      Fisher_p_value = NA
    )
  }
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
posthoc_results_fisher <- posthoc_results_fisher %>%
  mutate(p_value_adjusted = p.adjust(Fisher_p_value, method = "bonferroni"))
print(posthoc_results_fisher)
```
## Metric Analysis

__Figure S58: Boxplot of uni core length by site__
```{r}
uni_core_length_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length)),
  aes(x = reorder(Site, desc(Max_Length), FUN = median), y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S58-uni_core_length_site.png", plot = uni_core_length_site, width = 12, height = 8)
```
__Figure S59: Boxplot of uni core length by cluster__
```{r}
uni_core_length_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length)),
              aes(x = as.factor(Cluster_3), y = Max_Length,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max length (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S59-uni_core_length_cluster.png", plot = uni_core_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Test of uni core length__
```{r}
df_uni_L = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length))
kruskal.test(Max_Length ~ as.factor(Cluster_3), data = df_uni_L)
dunn.test(df_uni_L$Max_Length, df_uni_L$Cluster_3, method = "bonferroni")
```
__Figure S60: Boxplot of uni core width by site__
```{r}
uni_core_width_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Width)),
  aes(x = reorder(Site, desc(Max_Width), FUN = median), y = Max_Width, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S60-uni_core_width_site.png", plot = uni_core_width_site, width = 12, height = 8)
```
__Figure S61: Boxplot of uni core width by cluster__
```{r}
uni_core_width_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Width)),
              aes(x = as.factor(Cluster_3), y = Max_Width,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S61-uni_core_width_cluster.png", plot = uni_core_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of uni core width__
```{r}
df_uni_W = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Width))
kruskal.test(Max_Width ~ as.factor(Cluster_3), data = df_uni_W)
dunn.test(df_uni_W$Max_Width, df_uni_W$Cluster_3, method = "bonferroni")
```
__Figure S62: Boxplot of uni core thickness by site__
```{r}
uni_core_thick_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Thick)),
  aes(x = reorder(Site, desc(Max_Thick), FUN = median), y = Max_Thick, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickness (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_thick_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S62-uni_core_thick_site.png", plot = uni_core_thick_site, width = 12, height = 8)
```
__Figure S63: Boxplot of uni core thickness by cluster__
```{r}
uni_core_thick_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Thick)),
              aes(x = as.factor(Cluster_3), y = Max_Thick,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "midpoint thickness (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_thick_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S63-uni_core_thick_cluster.png", plot = uni_core_thick_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of uni core thickness__
```{r}
df_uni_T = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Thick))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_uni_T)
dunn.test(df_uni_T$Max_Thick, df_uni_T$Cluster_3, method = "bonferroni")
```
__Figure S64: Boxplot of uni core Index of Flattening by site__
```{r}
uni_core_IF_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_F)),
  aes(x = reorder(Site, desc(Index_F), FUN = median), y = Index_F, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_IF_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S64-uni_core_IF_site.png", plot = uni_core_IF_site, width = 12, height = 8)
```
__Figure S65: Boxplot of uni core Index of Flattening by cluster__
```{r}
uni_core_IF_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_F)),
              aes(x = as.factor(Cluster_3), y = Index_F,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "thickness / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_IF_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S65-uni_core_IF_cluster.png", plot = uni_core_IF_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's tests of uni core IF__
```{r}
df_uni_IF = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_F))
kruskal.test(Index_F ~ as.factor(Cluster_3), data = df_uni_IF)
dunn.test(df_uni_IF$Index_F, df_uni_IF$Cluster_3, method = "bonferroni")
```
__Figure S66: Boxplot of uni core Index of Elongation by site__
```{r}
uni_core_IE_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_E)),
  aes(x = reorder(Site, desc(Index_E), FUN = median), y = Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S66-uni_core_IE_site.png", plot = uni_core_IE_site, width = 12, height = 8)
```
__Figure S67: Boxplot of uni core Index of Elongation by cluster__
```{r}
uni_core_IE_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_E)),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S67-uni_core_IE_cluster.png", plot = uni_core_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of uni core IE__
```{r}
df_uni_IE = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Index_E))
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_uni_IE)
dunn.test(df_uni_IE$Index_E, df_uni_IE$Cluster_3, method = "bonferroni")
```
__Figure S68: uni core L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
     filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
uni_core_LW_scatter <- ggplot(
    df_filtered %>%  filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 0.9, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 35, y = y - 10, label = paste0("", round(estimate, 2), "")),
        hjust = 0.0, vjust = 0.0,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
uni_core_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S68-uni_core_LW_scatter.png", plot = uni_core_LW_scatter, width = 12, height = 8)
```
__Figure 13: Boxplot of uni core geometric mean by site__
```{r}
uni_core_geo_mean_site <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Geo_Mean)),
  aes(x = reorder(Site, desc(Geo_Mean), FUN = median), y = Geo_Mean, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
uni_core_geo_mean_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/13-uni_core_geo_mean_site.png", plot = uni_core_geo_mean_site, width = 12, height = 8)
```
__Figure S69: Boxplot of uni core geometric mean by cluster__
```{r}
uni_core_geo_mean_cluster <- ggplot(df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Geo_Mean)),
              aes(x = as.factor(Cluster_3), y = Geo_Mean,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "geometric mean (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
uni_core_geo_mean_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S69-uni_core_geo_mean_cluster.png", plot = uni_core_geo_mean_cluster, width = 12, height = 8)
```
__Kruskal Wallis & Dunn's Tests of uni core geo mean__
```{r}
df_UGM = df_filtered %>% filter(Art_Type %in% c("uni_blade", "uni_flake"), !is.na(Geo_Mean))
kruskal.test(Geo_Mean ~ as.factor(Cluster_3), data = df_UGM)
dunn.test(df_UGM$Geo_Mean, df_UGM$Cluster_3, method = "bonferroni")
```
# Debitage

## Metric Analysis

__Figure S70: debitage L x W scatterplot by site
```{r}
slopes <- df_filtered %>%
    filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Site) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Site, estimate)
label_positions <- df_filtered %>%
    filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Site) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Site")
debitage_LW_scatter_site <- ggplot(
    df_filtered %>% filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Site))
) +
    geom_point(size = 1.1, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 30, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Site",
    ) +
    scale_color_brewer(palette = "Set3") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
debitage_LW_scatter_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S70-debitage_LW_scatter_site.png", plot = debitage_LW_scatter_site, width = 12, height = 8)
```

__Figure S71: debitage L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
debitage_LW_scatter_cluster <- ggplot(
    df_filtered %>% filter(Art_Class == "debitage", !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 1.1, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 30, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_manual(values = c("I" = "#A6CEE3", "II" = "#1F78B4", "III" = "#B2DF8A")) +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
debitage_LW_scatter_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S71-debitage_LW_scatter_cluster.png", plot = debitage_LW_scatter_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of debitage IE__
```{r}
df_deb_IE = df_filtered %>% filter(Art_Class == "debitage", !is.na(Index_E))
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_deb_IE)
dunn.test(df_deb_IE$Index_E, df_deb_IE$Cluster_3, method = "bonferroni")
```
# Levallois Points

## Metric Analysis

### Lev Point Metrics

__Figure S72: Boxplot of Lev point length by site__
```{r}
lp_length_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Length)),
  aes(x = reorder(Site, desc(Max_Length), FUN = median), y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S72-lp_length_site.png", plot = lp_length_site, width = 12, height = 8)
```
__Figure S73: Boxplot of Lev point length by cluster__
```{r}
lp_length_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Length), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = Max_Length,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max length (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S73-lp_length_cluster.png", plot = lp_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's test of Lev point length__
```{r}
df_lp_L = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Length))
kruskal.test(Max_Length ~ as.factor(Cluster_3), data = df_lp_L)
dunn.test(df_lp_L$Max_Length, df_lp_L$Cluster_3, method = "bonferroni")
```
__Figure S74: Boxplot of Lev point width by site__
```{r}
lp_width_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Width)),
  aes(x = reorder(Site, desc(Max_Width), FUN = median), y = Max_Width, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S74-lp_width_site.png", plot = lp_width_site, width = 12, height = 8)
```
__Figure S75: Boxplot of Lev point width by cluster__
```{r}
lp_width_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Width), Art_Type_Sub == "point_Lev"),
  aes(x = as.factor(Cluster_3), y = Max_Width,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max width (mm)",
      fill="Cluster_3",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S75-lp_width_cluster.png", plot = lp_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point width__
```{r}
df_lp_W = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Width))
kruskal.test(Max_Width ~ as.factor(Cluster_3), data = df_lp_W)
```
__Figure S76: Boxplot of Lev point thickness by site__
```{r}
lp_thick_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Thick)),
  aes(x = reorder(Site, desc(Max_Thick), FUN = median), y = Max_Thick, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickness (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_thick_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S76-lp_thick_site.png", plot = lp_thick_site, width = 12, height = 8)
```
__Figure S77: Boxplot of Lev point thickness by cluster__
```{r}
lp_thick_cluster <- ggplot(df_filtered %>% filter(!is.na(Max_Thick), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = Max_Thick,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "midpoint thickness (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_thick_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S77-lp_thick_cluster.png", plot = lp_thick_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point thickness__
```{r}
df_lp_T = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Thick))
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_lp_T)
```
__Figure S78: Boxplot of Lev point Index of Flattening by site__
```{r}
lp_IF_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Index_F)),
  aes(x = reorder(Site, desc(Index_F), FUN = median), y = Index_F, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_IF_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S78-lp_IF_site.png", plot = lp_IF_site, width = 12, height = 8)
```
__Figure S79: Boxplot of Lev point Index of Flattening by cluster__
```{r}
lp_IF_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_F), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = Index_F,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "thickness / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_IF_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S79-lp_IF_cluster.png", plot = lp_IF_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point IF__
```{r}
df_lp_IF = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Index_F))
kruskal.test(Index_F ~ as.factor(Cluster_3), data = df_lp_IF)
```
__Figure S80: Boxplot of Lev point Index of Elongation by site__
```{r}
lp_IE_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Index_E)),
  aes(x = reorder(Site, desc(Index_E), FUN = median), y = Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S80-lp_IE_site.png", plot = lp_IE_site, width = 12, height = 8)
```
__Figure S81: Boxplot of Lev point Index of Elongation by cluster__
```{r}
lp_IE_cluster <- ggplot(df_filtered %>% filter(!is.na(Index_E), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S81-lp_IE_cluster.png", plot = lp_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Tests of Lev point IE__
```{r}
df_lp_IE = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Index_E))
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_lp_IE)
dunn.test(df_lp_IE$Index_E, df_lp_IE$Cluster_3, method = "bonferroni")
```
__Figure S82: Lev point L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Type_Sub == "point_Lev", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Max_Length ~ Max_Width, data = .))) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(Art_Type_Sub == "point_Lev", !is.na(Max_Length), !is.na(Max_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
lp_LW_scatter <- ggplot(
    df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Max_Length), !is.na(Max_Width)),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 0.9, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 13, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
lp_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S82-lp_LW_scatter.png", plot = lp_LW_scatter, width = 12, height = 8)
```
__Figure S83: Boxplot of Lev point geometric mean by site__
```{r}
lp_geo_mean_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Geo_Mean)),
  aes(x = reorder(Site, desc(Geo_Mean), FUN = median), y = Geo_Mean, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_geo_mean_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S83-lp_geo_mean_site.png", plot = lp_geo_mean_site, width = 12, height = 8)
```
__Figure 15: Boxplot of Lev point geometric mean by cluster__
```{r}
lp_geo_mean_cluster <- ggplot(df_filtered %>% filter(!is.na(Geo_Mean), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = Geo_Mean,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "geometric mean (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_geo_mean_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/15-lp_geo_mean_cluster.png", plot = lp_geo_mean_cluster, width = 12, height = 8)
```
__Kruskal Wallis & Dunn's tests of Lev point geo mean__
```{r}
df_lp_GM = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(Geo_Mean))
kruskal.test(Geo_Mean ~ as.factor(Cluster_3), data = df_lp_GM)
dunn.test(df_lp_GM$Geo_Mean, df_lp_GM$Cluster_3, method = "bonferroni")
```
### Tip Cross-Section

__Figure S84: Boxplot of Lev point TCSA by site__
```{r}
lp_TCSA_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(TCSA)),
  aes(x = reorder(Site, desc(TCSA), FUN = median), y = TCSA, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "TCSA (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_TCSA_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S84-lp_TCSA_site.png", plot = lp_TCSA_site, width = 12, height = 8)
```
__Figure S85: Boxplot of Lev point TCSA by cluster__
```{r}
lp_TCSA_cluster <- ggplot(df_filtered %>% filter(!is.na(TCSA), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = TCSA,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "TCSA (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_TCSA_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S85-lp_TCSA_cluster.png", plot = lp_TCSA_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point TCSA__
```{r}
df_TCSA = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(TCSA))
kruskal.test(TCSA ~ as.factor(Cluster_3), data = df_TCSA)
```
__Figure S86: Boxplot of Lev point TCSP by site__
```{r}
lp_TCSP_site <- ggplot(df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(TCSP)),
  aes(x = reorder(Site, desc(TCSP), FUN = median), y = TCSP, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "TCSP (mm)",
      fill = "Cluster") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_TCSP_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S86-lp_TCSP_site.png", plot = lp_TCSP_site, width = 12, height = 8)
```
__Figure S87: Boxplot of Lev point TCSP by cluster__
```{r}
lp_TCSP_cluster <- ggplot(df_filtered %>% filter(!is.na(TCSP), Art_Type_Sub == "point_Lev"),
              aes(x = as.factor(Cluster_3), y = TCSP,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "TCSP (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
lp_TCSP_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S87-lp_TCSP_cluster.png", plot = lp_TCSP_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point TCSP__
```{r}
df_TCSP = df_filtered %>% filter(Art_Type_Sub == "point_Lev", !is.na(TCSA))
kruskal.test(TCSP ~ as.factor(Cluster_3), data = df_TCSP)
```
### Lev Point Preferential Scar Metrics

__Figure S88: Boxplot of Lev scar length by site__
```{r}
ls_length_site <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Length)),
  aes(x = reorder(Site, desc(Lev_Scar_Length), FUN = median), y = Lev_Scar_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "scar length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ls_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S88-ls_length_site.png", plot = ls_length_site, width = 12, height = 8)
```
__Figure S89: Boxplot of Lev scar length by cluster__
```{r}
ls_length_cluster <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Length)),
              aes(x = as.factor(Cluster_3), y = Lev_Scar_Length,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "scar length (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ls_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S89-ls_length_cluster.png", plot = ls_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's test of Lev scar length__
```{r}
df_ls_L = df_filtered %>% filter(!is.na(Lev_Scar_Length))
kruskal.test(Lev_Scar_Length ~ as.factor(Cluster_3), data = df_ls_L)
dunn.test(df_ls_L$Lev_Scar_Length, df_ls_L$Cluster_3, method = "bonferroni")
```
__Figure S90: Boxplot of Lev scar width by site__
```{r}
ls_width_site <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Width)),
  aes(x = reorder(Site, desc(Lev_Scar_Width), FUN = median), y = Lev_Scar_Width,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "scar width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
lp_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S90-ls_width_site.png", plot = ls_width_site, width = 12, height = 8)
```
__Figure S91: Boxplot of Lev scar width by cluster__
```{r}
ls_width_cluster <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Width)),
  aes(x = as.factor(Cluster_3), y = Lev_Scar_Width,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "max width (mm)",
      fill="Cluster_3",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ls_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S91-ls_width_cluster.png", plot = ls_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point width__
```{r}
df_ls_W = df_filtered %>% filter(!is.na(Lev_Scar_Width))
kruskal.test(Lev_Scar_Width ~ as.factor(Cluster_3), data = df_ls_W)
dunn.test(df_ls_W$Lev_Scar_Width, df_ls_W$Cluster_3, method = "bonferroni")
```
__Figure S92: Boxplot of Lev point scar Index of Elongation by site__
```{r}
ls_IE_site <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Index_E)),
  aes(x = reorder(Site, desc(Lev_Scar_Index_E), FUN = median), y = Lev_Scar_Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ls_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S92-ls_IE_site.png", plot = ls_IE_site, width = 12, height = 8)
```
__Figure S93: Boxplot of Lev point scar Index of Elongation by cluster__
```{r}
ls_IE_cluster <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Index_E)),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length / width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ls_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S93-ls_IE_cluster.png", plot = ls_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Tests of Lev point scar IE__
```{r}
df_ls_IE = df_filtered %>% filter(!is.na(Lev_Scar_Index_E))
kruskal.test(Lev_Scar_Index_E ~ as.factor(Cluster_3), data = df_ls_IE)
dunn.test(df_ls_IE$Lev_Scar_Index_E, df_ls_IE$Cluster_3, method = "bonferroni")
```
__Figure S94: Lev point scar L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(!is.na(Lev_Scar_Length), !is.na(Lev_Scar_Width)) %>%
    group_by(Cluster_3) %>%
    do(tidy(lm(Lev_Scar_Length ~ Lev_Scar_Width, data = .))) %>%
    filter(term == "Lev_Scar_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(!is.na(Lev_Scar_Length), !is.na(Lev_Scar_Width)) %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Lev_Scar_Width),
        y = mean(Lev_Scar_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
ls_LW_scatter <- ggplot(
    df_filtered %>% filter(!is.na(Lev_Scar_Length), !is.na(Lev_Scar_Width)),
    aes(x = Lev_Scar_Width, y = Lev_Scar_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 0.9, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x + 15, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = 0, vjust = -0.5,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster",
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17),
    )
ls_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S94-ls_LW_scatter.png", plot = ls_LW_scatter, width = 12, height = 8)
```
__Figure S95: Boxplot of Lev point scar size by site__
```{r}
ls_s_site <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Size)),
  aes(x = reorder(Site, desc(Lev_Scar_Size), FUN = median), y = Lev_Scar_Size, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length x width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ls_s_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S95-ls_s_site.png", plot = ls_s_site, width = 12, height = 8)
```
__Figure S96: Boxplot of Lev point scar size by cluster__
```{r}
ls_s_cluster <- ggplot(df_filtered %>% filter(!is.na(Lev_Scar_Size)),
              aes(x = as.factor(Cluster_3), y = Index_E,
                  fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = .5) +
  labs(x = "", y = "length x width (mm)",
       fill="Cluster_3",
       title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ls_s_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S96-ls_s_cluster.png", plot = ls_s_cluster, width = 12, height = 8)
```
__Kruskal Wallis and Dunn's Tests of Lev point scar size__
```{r}
df_ls_s = df_filtered %>% filter(!is.na(Lev_Scar_Size))
kruskal.test(Lev_Scar_Size ~ as.factor(Cluster_3), data = df_ls_s)
dunn.test(df_ls_s$Lev_Scar_Size, df_ls_s$Cluster_3, method = "bonferroni")
```
### Comparison of Lev points versus Lev scars

__Figure S97: Boxplots of point vs. scar length by cluster__
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Length) & Art_Type == "Lev_Nub") | 
         (!is.na(Max_Length) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Length, Max_Length), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Length" ~ "core scar length",
    Variable == "Max_Length" ~ "Lev point length"
  ))
point_vs_scar_length <- ggplot(df_combined, aes(x = Variable, y = Value, fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length (mm)",
       fill = "Cluster_3",
       ) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 1, size = 13, angle = 45),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 13),
    strip.text = element_text(size = 17, face = "bold")
  ) +
  facet_grid(. ~ Cluster_3)
print(point_vs_scar_length)
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S97-point_vs_scar_length.png", plot = point_vs_scar_length, width = 12, height = 8)
```
__Kruskal Wallis test of Lev point versus scar length__ 
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Length) & Art_Type == "Lev_Nub") | 
         (!is.na(Max_Length) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Length, Max_Length), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Length" ~ "scar length",
    Variable == "Max_Length" ~ "Lev point length"
  ))

results <- df_combined %>%
  group_by(Cluster_3) %>%
  summarise(
    kruskal_test = list(kruskal.test(Value ~ Variable, data = cur_data()))
  )
results$kruskal_test
```
__Figure S98: Boxplots of point vs. scar width by cluster__
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Width) & Art_Type == "Lev_Nub") | 
         (!is.na(Max_Width) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Width, Max_Width), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Width" ~ "core scar width",
    Variable == "Max_Width" ~ "Lev point width"
  ))
point_vs_scar_width <- ggplot(df_combined, aes(x = Variable, y = Value, fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "width (mm)",
       fill = "Cluster_3",
       ) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 1, size = 13, angle = 45),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 13),
    strip.text = element_text(size = 17, face = "bold")
  ) +
  facet_grid(. ~ Cluster_3)
print(point_vs_scar_width)
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S98-point_vs_scar_width.png", plot = point_vs_scar_width, width = 12, height = 8)
```
__Kruskal Wallis tests of Lev point versus scar width__ 
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Width) & Art_Type == "Lev_Nub") | 
         (!is.na(Max_Width) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Width, Max_Width), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Width" ~ "scar width",
    Variable == "Max_Width" ~ "Lev point width"
  ))
results <- df_combined %>%
  group_by(Cluster_3) %>%
  summarise(
    kruskal_test = list(kruskal.test(Value ~ Variable, data = cur_data()))
  )
results$kruskal_test
```
__Figure 17: Boxplots of point vs. scar size by cluster__
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Size) & Art_Type == "Lev_Nub") | 
         (!is.na(Size) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Size, Size), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Size" ~ "core scar size",
    Variable == "Size" ~ "Lev point size"
  ))
point_vs_scar_size <- ggplot(df_combined, aes(x = Variable, y = Value, fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length x width (mm)",
       fill = "Cluster_3",
       ) +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 1, size = 13, angle = 45),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 13),
    strip.text = element_text(size = 17, face = "bold")
  ) +
  facet_grid(. ~ Cluster_3)
point_vs_scar_size
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/17-point_vs_scar_size.png", plot = point_vs_scar_size, width = 12, height = 8)
```
__Kruskal Wallis tests of Lev point versus scar size__
```{r}
df_combined <- df_filtered %>%
  filter((!is.na(Lev_Scar_Size) & Art_Type == "Lev_Nub") | 
         (!is.na(Size) & Art_Type_Sub == "point_Lev")) %>%
  pivot_longer(cols = c(Lev_Scar_Size, Size), names_to = "Variable", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  mutate(Variable = case_when(
    Variable == "Lev_Scar_Size" ~ "core scar size",
    Variable == "Size" ~ "Lev point size"
  ))
results <- df_combined %>%
  group_by(Cluster_3) %>%
  summarise(
    kruskal_test = list(kruskal.test(Value ~ Variable, data = cur_data()))
  )
results$kruskal_test
```
# Tools

## Types

__Data table of tool types by site__ 
```{r}
result <- df_filtered %>%
  filter(!is.na(Tool_Type)) %>%
  group_by(Site, Tool_Type) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Count_Percentage = paste0(Count, " (", round(Percentage, 2), "%)")) %>%
  select(Site, Tool_Type, Count_Percentage)
pivot_result <- result %>%
  arrange(Tool_Type) %>%
  pivot_wider(names_from = Site, values_from = Count_Percentage, values_fill = "-") %>%
  select(Tool_Type, order(as.numeric(gsub("[^0-9]", "", names(.)))))
pivot_result
```
__Table 3: Data table of tool types by cluster__ 
```{r}
result_cluster <- df_filtered %>%
  filter(!is.na(Tool_Type)) %>%
  group_by(Cluster_3, Tool_Type) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(Cluster_3) %>% 
  mutate(Percentage = Count / sum(Count) * 100) %>%
  mutate(Count_Percentage = paste0(Count, " (", round(Percentage, 1), "%)")) %>%
  select(Cluster_3, Tool_Type, Count_Percentage)
pivot_result_cluster <- result_cluster %>%
  arrange(Tool_Type) %>%
  pivot_wider(
    names_from = Cluster_3,
    values_from = Count_Percentage,
    values_fill = "-"
  ) %>%
  select(Tool_Type, `I`, `II`, `III`) 
pivot_result_cluster
```
__Figure S99: Bar plots of retouched tool counts per site (no Lev blanks)__
```{r}
df_ret_tool_counts <- df_filtered %>%
  filter(Art_Class == "tool", !is.na(Patination), Art_Type != "biface", Tool_Type != "Levallois_blank")
count_table <- df_ret_tool_counts %>%
  group_by(Site, Tool_Type) %>%
  summarize(count = n(), .groups = 'drop')
percentage_table <- count_table %>%
  group_by(Site) %>%
  mutate(percentage = (count / sum(count)) * 100)
ret_tool_counts_site <- ggplot(percentage_table, aes(x = factor(Site), y = percentage, fill = Tool_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_bar(stat = "identity", position = "stack") +
   geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(x = "", y = "Percentage", fill = "Tool Type") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
ret_tool_counts_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S99-ret_tool_counts_site.png", plot = ret_tool_counts_site, width = 12, height = 8)
```
__Figure S100: Bar plot of retouched tool counts per cluster__
```{r}
df_ret_tool_counts <- df_filtered %>%
  filter(Art_Class == "tool", !is.na(Patination), Art_Type != "biface", Tool_Type != "Levallois_blank")
count_table <- df_ret_tool_counts %>%
  group_by(Cluster_3, Tool_Type) %>%
  summarize(count = n(), .groups = 'drop')
percentage_table <- count_table %>%
  group_by(Cluster_3) %>%
  mutate(percentage = (count / sum(count)) * 100)
ret_tool_counts_cluster <- ggplot(percentage_table, aes(x = factor(Cluster_3), y = percentage, fill = Tool_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_bar(stat = "identity", position = "stack") +
   geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  labs(x = "", y = "Percentage", fill = "Tool Type") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
ret_tool_counts_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S100-ret_tool_counts_cluster.png", plot = ret_tool_counts_cluster, width = 12, height = 8)
```
## Blank Selection

__Figure S101: Bar plots of retouched tool blank selection count per site__
```{r}
df_toolblankcounts <- df_filtered %>%
  filter(Art_Class == "tool", Art_Type != "indeterm", Tool_Type != "Levallois_blank", Art_Type != "biface")
count_table <- df_toolblankcounts %>%
  group_by(Site, Art_Type) %>%
  summarize(count = n(), .groups = 'drop')
percentage_table <- count_table %>%
  group_by(Site) %>%
  mutate(percentage = (count / sum(count)) * 100)
tool_blank_counts_site <- ggplot(percentage_table, aes(x = factor(Site), y = percentage, fill = Art_Type)) +
  geom_bar(stat = "identity", position = "stack") +
   geom_text(
    aes(label = ifelse(percentage > 4.9, paste0(round(percentage, 1), "%\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3.5
  ) +
  labs(x = "", y = "Percentage", fill = "Blank Type") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 17, angle = 45, hjust = 1, face = "bold"),
    axis.title.y = element_text(size = 17),
    legend.position = "right",
    axis.text.y = element_text(size = 13),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 17)
  ) +
  scale_fill_brewer(palette = "Set3")
tool_blank_counts_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S101-tool_blank_counts_site.png", plot = tool_blank_counts_site, width = 12, height = 8)
```
__Figure 16: Pie charts of retouched tool blank selection by cluster__
```{r}
df_ret_tool_blank_counts <- df_filtered %>%
  filter(Art_Class == "tool", Art_Type != "indeterm", Tool_Type != "Levallois_blank", Art_Type != "biface")
data_counts <- df_ret_tool_blank_counts %>%
  group_by(Cluster_3, Art_Type) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_complete <- data_counts %>%
  complete(Cluster_3, Art_Type, fill = list(count = 0))
data_counts_complete <- data_counts_complete %>%
  group_by(Cluster_3) %>%
  mutate(proportion = count / sum(count))
ret_tool_blank_counts_cluster <- ggplot(data_counts_complete, aes(x = "", y = proportion, fill = Art_Type)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  facet_wrap(~ Cluster_3) +
  theme_void() +
 scale_fill_brewer(palette = "Set3")  +
  labs(title = "", fill = "") +
  geom_text(
    aes(label = ifelse(proportion > 0.049, paste0(scales::percent(proportion, accuracy = 1), "\n(", count, ")"), "")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme(
    plot.title = element_text(hjust = .5),
    legend.text = element_text(size = 13),   
    legend.position = "bottom",
    strip.text = element_text(size = 15, face = "bold")
  )
ret_tool_blank_counts_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-MS/16-ret_tool_blank_counts_cluster.png", plot = ret_tool_blank_counts_cluster, width = 12, height = 8)
```
__Chi Square Test of differences between clusters__
```{r}
df_ret_tool_blank_counts <- df_filtered %>%
  filter(Art_Class == "tool", Art_Type != "indeterm", Tool_Type != "Levallois_blank", Art_Type != "biface")
data_counts_tool <- df_ret_tool_blank_counts %>%
  group_by(Cluster_3, Art_Type) %>%
  summarize(count = n(), .groups = 'drop')
data_counts_tool_complete <- data_counts_tool %>%
  complete(Cluster_3, Art_Type, fill = list(count = 0))
contingency_table <- data_counts_tool_complete %>%
  select(Cluster_3, Art_Type, count) %>%
  spread(key = Art_Type, value = count, fill = 0) %>%
  column_to_rownames(var = "Cluster_3")
print(contingency_table)
chisq_test <- chisq.test(contingency_table)
chisq_test
```
__Fisher's Exact Test__
```{r}
clusters <- unique(df_ret_tool_blank_counts$Cluster_3)
cluster_pairs <- combn(clusters, 2, simplify = FALSE)
perform_fisher_posthoc <- function(pair) {
  cluster1 <- pair[1]
  cluster2 <- pair[2]
  subset_table <- data_counts_tool_complete %>%
    filter(Cluster_3 %in% c(cluster1, cluster2)) %>%
    select(Cluster_3, Art_Type, count) %>%
    spread(key = Art_Type, value = count, fill = 0) %>%
    column_to_rownames(var = "Cluster_3")
  if (nrow(subset_table) > 0 && ncol(subset_table) > 0) {
    fisher_result <- fisher.test(subset_table)
    p_value <- fisher_result$p.value
  } else {
    p_value <- NA
  }
  fisher_result <- fisher.test(subset_table)
  tibble(
    Cluster_1 = cluster1,
    Cluster_2 = cluster2,
    Fisher_p_value = fisher_result$p.value,
    p_value_adjusted = p.adjust(fisher_result$p.value, method = "bonferroni")
  )
}
posthoc_results_fisher <- purrr::map_df(cluster_pairs, perform_fisher_posthoc)
print(posthoc_results_fisher)
```
## Metric Analysis

__Figure S102: Boxplot of retouched tool length by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_length_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Max_Length), FUN = median), y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_length_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S102-ret_tool_length_site.png", plot = ret_tool_length_site, width = 12, height = 8)
```
__Figure S103: Boxplot of retouched tool length by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_length_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Max_Length, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max length (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_length_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S103-ret_tool_length_cluster.png", plot = ret_tool_length_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of retouched tool length__
```{r}
df_rt_L = df_filtered %>% filter(Art_Class == "tool", !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_rt_L)
```
__Figure S104: Boxplot of retouched tool width by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Width), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_width_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Max_Width), FUN = median), y = Max_Width, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_width_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S104-ret_tool_width_site.png", plot = ret_tool_width_site, width = 12, height = 8)
```
__Figure S105: Boxplot of retouched tool width by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Width), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_width_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Max_Width,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "max width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_width_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S105-ret_tool_width_cluster.png", plot = ret_tool_width_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of retouched tool width__
```{r}
df_rt_W = df_filtered %>% filter(Art_Class == "tool", !is.na(Max_Width), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Max_Width ~ as.factor(Cluster_3), data = df_rt_W)
dunn.test(df_rt_W$Max_Width, df_rt_W$Cluster_3, method = "bonferroni")
```
__Figure S106: Boxplot of retouched tool thickness by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Thick), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_thick_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Max_Thick), FUN = median), y = Max_Thick, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickess (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_thick_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S106-ret_tool_thick_site.png", plot = ret_tool_thick_site, width = 12, height = 8)
```
__Figure S107: Boxplot of retouched tool thickness by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Max_Thick), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_thick_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Max_Thick,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "midpoint thickness (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_thick_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S107-ret_tool_thick_cluster.png", plot = ret_tool_thick_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of retouched tool thickness__
```{r}
df_rt_T = df_filtered %>% filter(Art_Class == "tool", !is.na(Max_Thick), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Max_Thick ~ as.factor(Cluster_3), data = df_rt_T)
```
__Figure S108: Boxplot of retouched tool Index of Flattening by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Index_F), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_IF_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Index_F), FUN = median), y = Index_F, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_IF_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S108-ret_tool_IF_site.png", plot = ret_tool_IF_site, width = 12, height = 8)
```
__Figure S109: Boxplot of retouched tool Index of Flattening by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Index_F), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_IF_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Index_F,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "thickness / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_IF_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S109-ret_tool_IF_cluster.png", plot = ret_tool_IF_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of retouched tool IF__
```{r}
df_rt_IF = df_filtered %>% filter(Art_Class == "tool", !is.na(Max_Width), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Index_F ~ as.factor(Cluster_3), data = df_rt_IF)
```
__Figure S110: Boxplot of retouched tool Index of Elongation by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Index_E), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_IE_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Index_E), FUN = median), y = Index_E, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_IE_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S110-ret_tool_IE_site.png", plot = ret_tool_IE_site, width = 12, height = 8)
```
__Figure S111: Boxplot of retouched tool Index of Elongation by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Index_E), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_IE_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Index_E,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "length / width (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_IE_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S111-ret_tool_IE_cluster.png", plot = ret_tool_IE_cluster, width = 12, height = 8)
```
__Kruskal Wallis test of uni core IE__
```{r}
df_rt_IE = df_filtered %>% filter(Art_Class == "tool", !is.na(Index_E), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Index_E ~ as.factor(Cluster_3), data = df_rt_IE)
dunn.test(df_rt_IE$Index_E, df_rt_IE$Cluster_3, method = "bonferroni")
```
__Figure S112: retouched tool L x W scatterplot by cluster__
```{r}
slopes <- df_filtered %>%
    filter(Art_Class == "tool", !is.na(Max_Width), !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface") %>%
    group_by(Cluster_3) %>%
    nest() %>%
    mutate(model = map(data, ~ lm(Max_Length ~ Max_Width, data = .x)),
           tidied = map(model, tidy)) %>%
    unnest(tidied) %>%
    filter(term == "Max_Width") %>%
    select(Cluster_3, estimate)
label_positions <- df_filtered %>%
    filter(Art_Class == "tool", !is.na(Max_Width), !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface") %>%
    group_by(Cluster_3) %>%
    summarize(
        x = mean(Max_Width),
        y = mean(Max_Length)
    ) %>%
    left_join(slopes, by = "Cluster_3")
ret_tool_LW_scatter <- ggplot(
    df_filtered %>% filter(Art_Class == "tool", !is.na(Max_Width), !is.na(Max_Length), Tool_Type != "Levallois_blank", Art_Type != "biface"),
    aes(x = Max_Width, y = Max_Length, color = as.factor(Cluster_3))
) +
    geom_point(size = 1.1, alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 1.2) +
    geom_text(
        data = label_positions,
        aes(x = x, y = y, label = paste0("", round(estimate, 2), "")),
        hjust = -5.2, vjust = 0.0,
        show.legend = FALSE
    ) +
    labs(
        x = "width (mm)",
        y = "length (mm)",
        color = "Cluster"
    ) +
    scale_color_brewer(palette = "Paired") +
    theme_minimal() +
    theme(
        legend.title = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17),
        axis.title.x = element_text(size = 17)
    )
ret_tool_LW_scatter
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S112-ret_tool_LW_scatter.png", plot = ret_tool_LW_scatter, width = 12, height = 8)
```
__Figure S113: Boxplot of retouched tool geometric mean by site__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Geo_Mean), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_GM_site <- ggplot(filtered_data, aes(x = reorder(Site, desc(Geo_Mean), FUN = median), y = Geo_Mean, 
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 17),
    legend.text = element_text(size = 13),
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 17, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.title.y = element_text(size = 17)
    )
ret_tool_GM_site
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S113-ret_tool_GM_site.png", plot = ret_tool_GM_site, width = 12, height = 8)
```
__Figure S114: Boxplot of retouched tool geometric mean by cluster__
```{r}
filtered_data <- df_filtered %>% 
  filter(Art_Class == "tool", !is.na(Geo_Mean), Tool_Type != "Levallois_blank", Art_Type != "biface")
ret_tool_GM_cluster <- ggplot(filtered_data, aes(x = Cluster_3, y = Geo_Mean,
      fill = as.factor(Cluster_3))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 0.5) +
  labs(x = "", y = "geometric mean (mm)",
      fill = "Cluster",
      title = "") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  theme(
    legend.position = "none",
        axis.text.x = element_text(hjust = 1, size = 17, face = "bold"),
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 17)
    )
ret_tool_GM_cluster
ggsave("~/Dropbox/Documents/My Publications/Nubian Tradition/Figures/R-SI/S114-ret_tool_GM_cluster.png", plot = ret_tool_GM_cluster, width = 12, height = 8)
```
__Kruskal Wallis & Dunn's Tests of retouched tool geo mean__
```{r}
df_rt_GM = df_filtered %>%  filter(Art_Class == "tool", !is.na(Geo_Mean), Tool_Type != "Levallois_blank", Art_Type != "biface")
kruskal.test(Geo_Mean ~ as.factor(Cluster_3), data = df_rt_GM)
dunn.test(df_rt_GM$Geo_Mean, df_rt_GM$Cluster_3, method = "bonferroni")
```